---
title: "WiDS_Datathon_2022"
author: "Uyen Nguyen"
date: "2/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
library(tidyverse)
library(corrplot)
library(stringr)
library(ggcorrplot)

test <- tibble(read.csv('test.csv'))
train <- tibble(read.csv('train.csv'))
```


### EDA

```{r}
# Making two vectors to contain column names of train dataset and how many NAs
# there are for each column

na_col <- c()
col_names <- colnames(train)

for (i in 1:ncol(train)) {
  na_col[i] <- sum(is.na(train[,i]))
}

tibble(col_names, na_col) %>%
  filter(na_col != 0) %>%  
  arrange(desc(na_col))
```

The variables:

* days_with_fog
* direction_peak_wind_speed
* direction_max_wind_speed
* max_wind_speed 

all have NAs for more than half of the rows in train. While

* energy_star_rating
* year_built 

have NAs for less than half of the rows in train. There are **`r sum(complete.cases(train))` (less than a quarter)** complete cases in train. This might be something to keep in mind 
regarding the approach we will take.  

Since we're not entirely sure if we'll be using the train dataset that includes NAs
in several columns or we'll be using only the completed cases instead. I created 
the train_complete_cases tibble with only complete cases free from columns' NAs
mentioned above. 

(If we decide to use train_complete_cases to fit our model, we should also bootstrap
our training tibble since we took a massive cut to the data set.)

```{r}
train_complete_cases <- train[complete.cases(train),]
summary(train_complete_cases)
```

After gathering only the complete cases, I noticed that days_above_110F had 0
for both min and max so I took that column out. 

```{r}
train_complete_cases <- train_complete_cases %>% select(-days_above_110F)
```

The correlation matrix is produced below. 

```{r out.width='100%', out.height='100%', fig.align='center'}
ggcorrplot(cor(train_complete_cases %>% select(where(is.numeric))), type = 'lower', tl.cex = 5, ggtheme = ggplot2::theme_void())
```

There are no real variables that jumped out for the response variable site_eui 
from the correlation matrix so we'll proceed with caution on any linear assumptions.

For the sake of visualization, I grouped values in the variable facility_type into 
broader categories such as Public, Education, Healthcare, etc. to better visualize
these groups. 

```{r}
# Making a function to convert details in facility_type into broad groups in
# variable class
train_complete_cases$class <- ''

train_complete_cases[train_complete_cases$facility_type %in% c('2to4_Unit_Building', '5plus_Unit_Building', 
                                 'Mixed_Use_Commercial_and_Residential', 
                                 'Mixed_Use_Predominantly_Commercial', 
                                 'Mixed_Use_Predominantly_Residential', 
                                 'Multifamily_Uncategorized'),]$class <- 'Residential'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Public'),]$class <- 'Public'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Education'),]$class <- 'Education'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Food Service'),]$class <- 'Food Service'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Food'),]$class <- 'Food Service'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Health'),]$class <- 'Health Care'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Office'),]$class <- 'Office'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Lodging'),]$class <- 'Lodging'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Nursing'),]$class <- 'Health Care'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Retail'),]$class <- 'Retail'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Warehouse'),]$class <- 'Warehouse'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Service'),]$class <- 'Service'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Religious'),]$class <- 'Religious'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Industrial'),]$class <- 'Industrial'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Laboratory'),]$class <- 'Health Care'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Commercial'),]$class <- 'Commercial'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Data'),]$class <- 'Data Center'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Grocery'),]$class <- 'Grocery'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Parking'),]$class <- 'Parking Garage'
train_complete_cases[str_which(train_complete_cases$facility_type, 'Religious'),]$class <- 'Religious'
```
Here are some plots to visualize this data. 

```{r}
ggplot(train_complete_cases) +
  geom_point(aes(x=floor_area, y=site_eui, color = energy_star_rating)) +
  facet_wrap(~class)

ggplot(train_complete_cases, aes(x=site_eui, fill = class)) +
  geom_histogram()

ggplot(train_complete_cases, aes(x=energy_star_rating, fill = class)) +
  geom_histogram()

ggplot(train_complete_cases, aes(x=avg_temp, fill = class)) +
  geom_histogram()
```

## Data Analysis



```{r}
set.seed(1)

### Train data set provided

# na_train_1 <- lm(site_eui ~ year_built + floor_area + cooling_degree_days + heating_degree_days +
#    facility_type + avg_temp, data = train)
# sum_train_1 <- summary(na_train_1)
# sqrt(mean(sum_train_1$residuals^2))
# 
# yhat<-na_train_1$fitted.values
# res<-sum_train_1$residuals
# Data<-data.frame(yhat,res)
# 
# ggplot(Data, aes(x=yhat,y=res))+
# geom_point()+
# geom_hline(yintercept=0, color="red")+
# labs(x="Fitted y",
# y="Residuals",
# title="Residual Plot")

na_train_2 <- lm(site_eui ~ Year_Factor + floor_area + year_built + energy_star_rating, data = train)
sum_train_2 <-summary(na_train_2)
sqrt(mean(sum_train_2$residuals^2))

yhat<-na_train_2$fitted.values
res<-sum_train_2$residuals
Data<-data.frame(yhat,res)

ggplot(Data, aes(x=yhat,y=res))+
geom_point()+
geom_hline(yintercept=0, color="red")+
labs(x="Fitted y",
y="Residuals",
title="Residual Plot")

# na_train_3 <- lm(site_eui ~ Year_Factor + floor_area + year_built + energy_star_rating + avg_temp + heating_degree_days + march_avg_temp + june_avg_temp + august_avg_temp + 
# november_avg_temp + december_avg_temp, data = train)
# sum_train_3 <-summary(na_train_3)
# sqrt(mean(sum_train_3$residuals^2))
# 
# yhat<-na_train_3$fitted.values
# res<-sum_train_3$residuals
# Data<-data.frame(yhat,res)
# 
# ggplot(Data, aes(x=yhat,y=res))+
# geom_point()+
# geom_hline(yintercept=0, color="red")+
# labs(x="Fitted y",
# y="Residuals",
# title="Residual Plot")

### Complete cases of train data set

# comp_train_1 <- lm(site_eui ~ year_built + floor_area + cooling_degree_days + heating_degree_days +
#    facility_type + avg_temp, data = train_complete_cases)
# sum_train_1_comp <- summary(comp_train_1)
# sqrt(mean(sum_train_1_comp$residuals^2))
# 
# yhat<-comp_train_1$fitted.values
# res<-sum_train_1_comp$residuals
# Data<-data.frame(yhat,res)
# 
# ggplot(Data, aes(x=yhat,y=res))+
# geom_point()+
# geom_hline(yintercept=0, color="red")+
# labs(x="Fitted y",
# y="Residuals",
# title="Residual Plot")
# 
# comp_train_2 <- lm(site_eui ~ Year_Factor + floor_area + year_built + energy_star_rating, data = train_complete_cases)
# sum_train_2_comp <-summary(comp_train_2)
# sqrt(mean(sum_train_2_comp$residuals^2))
# 
# yhat<-comp_train_2$fitted.values
# res<-sum_train_2_comp$residuals
# Data<-data.frame(yhat,res)
# 
# ggplot(Data, aes(x=yhat,y=res))+
# geom_point()+
# geom_hline(yintercept=0, color="red")+
# labs(x="Fitted y",
# y="Residuals",
# title="Residual Plot")
# 
# comp_train_3 <- lm(site_eui ~ Year_Factor + floor_area + year_built + energy_star_rating + avg_temp + heating_degree_days + march_avg_temp + june_avg_temp + august_avg_temp + 
# november_avg_temp + december_avg_temp, data = train_complete_cases)
# sum_train_3_comp <-summary(comp_train_3)
# sqrt(mean(sum_train_3_comp$residuals^2))
# 
# yhat<-comp_train_3$fitted.values
# res<-sum_train_3_comp$residuals
# Data<-data.frame(yhat,res)
# 
# ggplot(Data, aes(x=yhat,y=res))+
# geom_point()+
# geom_hline(yintercept=0, color="red")+
# labs(x="Fitted y",
# y="Residuals",
# title="Residual Plot")
```